Interpretable Machine Learning

### 2018-10-19 poster/poster_2018.Rnw
Improves poster

\includegraphics[width=0.25\textwidth]{figure/black-box.png}

@@

\block{Why explain?}{
We need to explain machine learning models to\cite{adadi2018peeking}:
TODO: COPY FROM SLIDES
\begin{itemize}
\item \textbf{Explain to justify}: Machine learning models make errors, can be unfair (e.g racial bias.) and make unexpected predictions. In those cases, we require a model to justify predictions.
\item \textbf{Explain to control}: Understand more about the systems vulnerabilities and conditions for failure. e.g. know for what type of data it performs worse. 
\item \textbf{Explain to improve}: Insights about model behaviour helps to improve the model.
\item \textbf{Explain to discover}: A black box model is optimized to make predictions, not to give insights.
Interpretability methods enable to extract insights from black box models.
\end{itemize}
}

\block{Feature importance}{
We can measure importance of the inputs as the drop in performance of the model if we would not know the features.
Not knowing the feature can be simulated by shuffling the feature in the data: We break the assocation between the feature and the target.
If the performance of the model drops a lot, the features was important.
<<feature-importance, echo=FALSE>>=
imp = FeatureImp$new(pred, "mae")
plot(imp)
@
In our little rent example, the most important feature was the size of the living area.

Caveats: Doesn't work well when the features are highly correlated.
}

@@

<<ice, echo=FALSE>>=
eff = FeatureEffect$new(pred, "size", method = "pdp+ice")
plot(eff, rug = FALSE)  + ggtitle("Effect of size on predicted rent") +
  scale_x_continuous("Size of appartment") + 
  scale_y_continuous("Predicted rent")
@

@@

<<appartment, echo=FALSE,include=FALSE>>=
# Choose appartment
appartment = df[10,]
pred.rent = round(pred$predict(appartment)$.prediction, 0)
@

We have a couple of options to explain why a particular prediction was made by a machine learning model.
Suppose the rent model predicts \Sexpr{pred.rent} Euros rent, for an appartment in the inner city with a living area of 85 $m^2$ and where cats are allowed. 
How can we explain this prediction?

<<counterfact, echo=FALSE, include = FALSE>>=
appartment.mod = appartment
appartment.mod$cat.allowed = factor("No", levels = c("No", "Yes"))
pred.new = round(pred$predict(appartment.mod[,1:3, drop=FALSE]), 0)
@
\textbf{Counterfactual explanations}: 
A counterfactual explanation of a prediction describes the smallest change to the feature values (inputs) that changes the prediction to a predefined output.
For example: How does the input have to change so that the prediction changes below 1200 Euros?
Answer: If cats were not allowed, the predicted rent would be \Sexpr{pred.new} Euros.

@@

<<shapley, echo=FALSE>>=
shap = Shapley$new(pred)
shap$explain(appartment[,1:3])
plot(shap)
@

@@

\block{Interpretable Models}{
Some models are intrinsically interpretable, because they learn simple relationships between input and output like a decision tree or a weighted sum of the inputs.
The big disadvantage of intrinsically interpretable models is that they often predict the data less well.

@@

$x_1$ to $x_p$ are the p features that are used to predict the outcome y. 
The goal is to find the best possible weights, which can be found for example via the least squares algorithm.

In our rent example, the folowing weights are found:
<<train-model-lm, echo=FALSE, include=FALSE>>=
mod.lm = lm(rent ~ ., data = df)
cf = round(coef(mod.lm))
knitr::kable(cf, format = 'latex')
@

$$rent = \Sexpr{cf["(Intercept)"]} + \Sexpr{cf["size"]} \cdot size  \Sexpr{cf["locationOuter city"]} \cdot I_{\text{outer city}} +  \Sexpr{cf["cat.allowedYes"]} \cdot I_{\text{cats allowed}}$$

@@

\section*{Decision rules}
A decision rule is a simple IF-THEN statement consisting of a condition and a prediction.
For example: IF it rains today AND if it’s April (condition), THEN it will rain tomorrow (prediction).
A single decision rule or a combination of several rules can be used to make predictions.

@@

Other approaches: 
RuleFit CITE, SLIM, \ldots

>>>>

\includegraphics[width=0.25\textwidth]{figure/teaser.png}

@@

\block{Why explain?}{
We need to explain machine learning models to \cite{adadi2018peeking}:
\begin{itemize}
\item \textbf{Explain to justify}: Machine learning models make errors, can be unfair (e.g racial bias.) and make unexpected predictions. In those cases, we require a model to justify predictions.
\item \textbf{Explain to control}: Understand more about the systems vulnerabilities and conditions for failure. e.g. know for what type of data it performs worse. 
\item \textbf{Explain to improve}: Insights about model behaviour helps to improve the model.
\item \textbf{Explain to discover}: A black box model is optimized to make predictions, not to give insights.
Interpretability methods enable to extract insights from black box models.
\end{itemize}
}

\block{Feature importance}{
We can measure importance of the inputs as the drop in performance of the model if we would not know the features.
Not knowing the feature can be simulated by shuffling the feature in the data: We break the assocation between the feature and the target.
If the performance of the model drops a lot, the features was important.
<<feature-importance, echo=FALSE, fig.width=12, fig.height=3>>=
imp = FeatureImp$new(pred, "mae")

impres = imp$results

ggplot(impres) + geom_col(aes(y = importance, x = feature)) + coord_flip() +
  theme(aspect.ratio=1/4, text = element_text(size=20),
    panel.grid.major = element_blank(), panel.grid.minor = element_blank())
@
In our little rent example, the most important feature was the size of the living area. 
}

@@

<<ice, echo=FALSE, fig.width=12, fig.height=3>>=
eff = FeatureEffect$new(pred, "size", method = "pdp+ice")
plot(eff, rug = FALSE)  + ggtitle("Effect of size on predicted rent") +
  scale_x_continuous("Size of appartment") + 
  scale_y_continuous("Predicted rent") +
  theme(aspect.ratio=1/4, text = element_text(size=20),
    panel.grid.major = element_blank(), panel.grid.minor = element_blank())

@

@@

<<appartment, echo=FALSE,include=FALSE>>=
# Choose appartment
appartment = df[3,]
pred.rent = round(pred$predict(appartment)$.prediction, 0)
@

We have a couple of options to explain why a particular prediction was made by a machine learning model.
Suppose the rent model predicts \Sexpr{pred.rent} Euros rent, for an appartment in the outer city with a living area of 92 $m^2$ and where cats are not allowed. 
How can we explain this prediction?

<<counterfact, echo=FALSE, include = FALSE>>=
appartment.mod = appartment
appartment.mod$cat.allowed = factor("Yes", levels = c("No", "Yes"))
pred.new = round(pred$predict(appartment.mod[,1:3, drop=FALSE]), 0)
@
\textbf{Counterfactual explanations}: 
A counterfactual explanation of a prediction describes the smallest change to the feature values (inputs) that changes the prediction to a predefined output.
For example: How does the input have to change so that the prediction changes above 1100 Euros?
Answer (one of many possible ones): If cats were allowed, the predicted rent would be \Sexpr{pred.new} Euros.

@@

<<shapley, echo=FALSE, fig.width=12, fig.height=3>>=
shap = Shapley$new(pred)
shap$explain(appartment[,1:3])
plot(shap) +
  theme(aspect.ratio=1/4, text = element_text(size=20),
    panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  scale_y_continuous("Contribution to prediction") +
  scale_x_discrete("")

@

@@

\block{Interpretable Models}{
Some models are intrinsically interpretable, because they learn simple relationships between input and output.
Their big disadvantage is that they often predict the data less well.

@@

$x_1$ to $x_p$ are the features that are used to predict the outcome y. 
The goal is to find the best possible weights $\beta$.
In our rent example, the best weights are:
<<train-model-lm, echo=FALSE, include=FALSE>>=
mod.lm = lm(rent ~ ., data = df)
cf = round(coef(mod.lm))
knitr::kable(cf, format = 'latex')
@

$$\text{rent} = \Sexpr{cf["(Intercept)"]} + \Sexpr{cf["size"]} \cdot \text{size}  \Sexpr{cf["locationOuter city"]} \cdot I_{\text{outer city}} +  \Sexpr{cf["cat.allowedYes"]} \cdot I_{\text{cats allowed}}$$

@@

\section*{Decision Rules}
A decision rule is a simple IF-THEN statement consisting of a condition and a prediction.
For example: IF it rains today AND if it’s April (condition), THEN it will rain tomorrow (prediction).
A single decision rule or a combination of several rules can be used to make predictions.

@@
